---
title: "Predicting attitudes towards racist college professors"
author: "Ojaswi Malik"
date: "`r lubridate::today()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rcfss)
library(modelr)
library(broom)
library(knitr)
theme_set(theme_minimal())
```

```{r data}
#loading the data
gss_data <- rcfss::gss_colrac

#convert colrac to binary
gss_data <- gss_data %>%
  mutate(colrac = as.numeric(colrac))

```

## Using only Age to make prediction

In this analysis I will estimate a logistic regression model of whether or not an individual believes a racist professor should be allowed to teach with 'age' as the predictor. I will estimate predicted probabilities and plot the logistic regression line using ggplot.

```{r base age}
#estimating the model
age_model <- glm(colrac ~ age, data = gss_data,
                family = binomial)

#present as a tidy table
tidy(age_model) %>%
  kable()

#estimating the predicted probabilities
age_data <- augment(age_model,
                     newdata = data_grid(gss_data, age),
                     type.predict = "response")

#graphing the line
ggplot(age_data, aes(age, .fitted)) +
  geom_line() +
  labs(title = "Relationship between age and attitudes towards racist college professors",
       y = "Predicted Probability of allowing racist prof. to teach",
       x = "Age") +
  scale_y_continuous(limits = c(0, 1))

# Predicting the accuracy of the above model
age_model_accuracy <- augment(age_model, type.predict = "response") %>%
  mutate(.pred = as.numeric(.fitted > .5))
    
(age_model_err <- mean(age_model_accuracy$colrac != age_model_accuracy$.pred,
                      na.rm = TRUE))

```

**Observation**: From the above graph, we can observe that there is a negative correlation between age and attitude towards allowing racist professors to teach. This means that older individuals believe there is less probability for racist professors to teach. 

Although the p-value > 0.05 making it statistically significant, the significance of this correlation is still quite low. Moreover, The above model  has an error rate of 49.3%. This is a very high error rate and hence we need to construct other models.

## Using respondent's age and race (African American) as interactive predictors

```{r age and race}
#estimating the model
age_black_model <- glm(colrac ~ age * black, data = gss_data,
                family = binomial)

#present as a tidy table
tidy(age_black_model) %>%
  mutate(term = recode(
    term,
    age = "Age",
    blackYes = "Black",
    "age:blackYes" = "Age * Black"
  )) %>%
  kable()

#estimating the predicted probabilities
age_black_data <- augment(age_black_model,
                     newdata = data_grid(gss_data, age, black),
                     type.predict = "response")

#graphing the line
ggplot(age_black_data, aes(age, .fitted, color = black)) +
  geom_line() +
  labs(title = "Relationship between age and attitudes towards racist college professors",
       subtitle = "Interactive",
       y = "Predicted Probability of allowing racist prof. to teach",
       x = "Age",
       color = "If race is black:") +
  scale_y_continuous(limits = c(0, 1))

#calculating the error
age_black_model_accuracy <- augment(age_black_model, type.predict = "response") %>%
  mutate(.pred = as.numeric(.fitted > .5))
    
(age_black_model_err <- mean(age_black_model_accuracy$colrac != age_black_model_accuracy$.pred,
                      na.rm = TRUE))

```

**Observation**: This model shows that young respondents who are black believe that racist professors have a less probability of being allowed to teach. This is contrast with young respondents who are not black. However this gap in different opinions reduces with increase in age. Hence, old respondents at the age of ~90 have the same attitude towards racist professors.

This may be explained by young African Americans' lower tolerance towards racist professors and demanding a change, as compared to their older counterparts somewhat accepting this racial bias as a bitter reality. However, the error rate of this model is 44.8%. Although lower than the previous model, there is still room for improvement.

## Using respondent's age and social conservatism as interactive predictors

```{r}
#estimating the model
age_soc_model <- glm(colrac ~ age * social_cons3, data = gss_data,
                family = binomial)

#present as a tidy table
tidy(age_soc_model) %>%
  mutate(term = recode(
    term,
    age = "Respondent Age",
    social_cons3Mod = "Moderate Social Conservatism",
    social_cons3Conserv = "Conservative Social Conservatism",
    "age:social_cons3Mod" = "Age * Moderate Social Conservatism",
    "age:social_cons3Conserv" = "Age * Conservative Social Conservatism"
  )
) %>%
  kable()

#estimating the predicted probabilities
age_soc_data <- augment(age_soc_model,
                     newdata = data_grid(gss_data, age, social_cons3),
                     type.predict = "response")

#graphing the line
ggplot(age_soc_data, aes(age, .fitted, color = social_cons3)) +
  geom_line() +
  labs(title = "Relationship between age and attitudes towards racist college professors",
       y = "Predicted Probability of allowing racist prof. to teach",
       subtitle = "Interactive",
       x = "Age",
       color = "Social Conservatism Identification:") +
  scale_y_continuous(limits = c(0, 1))

#calculating the error
age_soc_model_accuracy <- augment(age_soc_model, type.predict = "response") %>%
  mutate(.pred = as.numeric(.fitted > .5))
    
(age_soc_model_err <- mean(age_soc_model_accuracy$colrac != age_soc_model_accuracy$.pred,
                      na.rm = TRUE))

```

**Observation**: From the above graph, we can ascertain that younger Moderates and Conservatives have the belief that racist professors have a less probability of being allowed to teach when compared to Liberals of the same age. This 







## Using authoritarianism and geographic location as independant variables to make prediction

```{r}
#estimating the model
auth_geo_model <- glm(colrac ~ authoritarianism + south, data = gss_data,
                family = binomial)

#present as a tidy table
tidy(auth_geo_model) %>%
  kable()

#estimating the predicted probabilities
auth_geo_data <- augment(auth_geo_model,
                     newdata = data_grid(gss_data, authoritarianism, south),
                     type.predict = "response")

#graphing the line
ggplot(auth_geo_data, aes(authoritarianism, .fitted, color = south)) +
  geom_line() +
  labs(title = "Impact of authoritarianism and location in attitudes towards racist college professors",
       subtitle = "Independantly",
       y = "Predicted Probability of allowing racist prof. to teach",
       x = "Authoritarianism Scale") +
  scale_y_continuous(limits = c(0, 1))

# Predicting the accuracy of the above model
auth_geo_model_accuracy <- augment(auth_geo_model, type.predict = "response") %>%
  mutate(.pred = as.numeric(.fitted > .5))
    
(auth_geo_model_err <- mean(auth_geo_model_accuracy$colrac != auth_geo_model_accuracy$.pred,
                      na.rm = TRUE))



```

**Observation**: There is a negative correlation as depicted from the negative coefficient estimate between authoritarianism and attitudes towards racist professors. Hence, respondents who are less authoritarian believe racist professors have a higher chance of being allowed to teach. However, this result is not statistically significant as the p-value is less than 5% 

People who belong from the south believe that there is less probability of allowing racist professors to teach as compared to those who do not belong from the south. The p-value for south is 0.309 which is greater than 0.05, making it statistically significant. 

However, this model has an error of 44.8%. Although lower than before, we should construct other models too.

## Using authoritarianism and geographic location as interactive variables to make prediction

In the previous model, I looked at authoritarianism and geographic location independantly. In this model, I want to see if there is any difference from these two variables interacting.


```{r}
#estimating the interactive model
auth_geo_model_interactive <- glm(colrac ~ authoritarianism * south, data = gss_data,
                family = binomial)

#present as a tidy table
tidy(auth_geo_model_interactive) %>%
  mutate(term = recode(
    term,
    "authoritarianism:southSouth" = "Interaction: Authoritarianism and South"
  )) %>%
  kable() 

#estimating the predicted probabilities
auth_geo_data_interactive <- augment(auth_geo_model_interactive,
                     newdata = data_grid(gss_data, authoritarianism, south),
                     type.predict = "response")

#graphing the line
ggplot(auth_geo_data_interactive, aes(authoritarianism, .fitted, color = south)) +
  geom_line() +
  labs(title = "Impact of authoritarianism and location in attitudes towards racist college professors",
       subtitle = "Interactive",
       y = "Predicted Probability of allowing racist prof. to teach",
       x = "Authoritarianism Scale") +
  scale_y_continuous(limits = c(0, 1))

```

**Observation**: In this model, the p-value for the interaction between Authoritarianism and South is 0.45 which makes statistically significant. There is a negative coefficient indicating the negative relationship observed the graph.
Interestingly, different from the above graph, respondents with very low authoritarianism from the south believe racist professors have a higher chance of being allowed to teach. Otherwise the results are similiar to the previous model, only that the gap between the Southerners and Non-Southerns increases as authoritarianism of respondents increases.

## Plotting Average Marginal Effects

Plotting the average marginal effect of each variable using the margins package to visualize the logistic regression estimates and confidence intervals.





Plot the average marginal effect of each variable using the margins package to visualize the logistic regression estimates and confidence intervals.## Session info

```{r}
devtools::session_info()
```
